Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
done with Section 5.1 Analysis

When considering a single proposal $X_i$ in isolation, we wish to compute:
\begin{align}\label{eq1}
\Pr(\alpha^*_i > T_i \mid X_i),
\end{align}
where
\begin{itemize}
\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
(e.g., $f=\frac{1}{2}$ for a simple majority).
\end{itemize}
Using a desirable prior
\begin{align}\label{eq2}
p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
\end{align}
corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
\begin{align}\label{eq3}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
\end{align}
for
\begin{align*}
&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
\end{align*}
where
\begin{itemize}
\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.,
\item[$c = $] $10$, a hyperparameter chosen by the community based on their relative concern about false alarms versus (temporarily) missed detection.
\end{itemize}

@@

Using Eqs.~\eqref{eq2} and \eqref{eq4}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.  For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.

@@

To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine two toy scenarios.

In Figure \ref{fig:credible_interval}, we examine how the $90\%$ credible interval converges as more votes are collected out of a population of $100$.  A statistical quorum is achieved when the credible interval is contained entirely above the x-axis.  In this example, the latent bias of the community is that $75\%$ of the community supports the proposal while $25\%$ opposes it.

In
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{figures/credible_interval.pdf}
\caption{.}
\label{fig:credible_interval}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=.45\textwidth]{figures/simple_majority.pdf}
\caption{.}
\label{fig:simple_majority}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=.45\textwidth]{figures/competing_proposals.pdf}
\caption{.}
\label{fig:simple_majority}
\end{figure}

>>>>

When considering a single proposal $X_i$ in isolation, we wish to compute:
\begin{align}\label{eq1}
\Pr(\alpha^*_i > T_i \mid X_i),
\end{align}
where
\begin{itemize}
\item[$\alpha^*_i = $] the total number of ``yes'' votes for $X_i$ if we were to sample the entire population,
\item[$T_i = $] $f \cdot N_i$, for some fraction $f$ corresponding to the type of majority needed for the vote
(e.g., $f=\frac{1}{2}$ for a simple majority).
\end{itemize}
Using the desirable prior
\begin{align}\label{eq2}
p_i \mid X_i \sim \operatorname{Beta}(1+\frac{\alpha_i}{K_i},1+\frac{\beta_i}{K_i}),
\end{align}
corresponding to a uniform prior when no votes have been cast ($\alpha_i = \beta_i = 0$), we can express \eqref{eq1} as:
\begin{align}\label{eq3}
\Pr(\alpha^*_i > T_i \mid X_i) &= 1 - {CDF}_{\nu_i}(V_i)
\end{align}
for
\begin{align*}
&\nu_i \sim \operatorname{Beta-Binomial}(1 + \frac{\alpha_i}{K_i}c, 1 + \frac{\beta_i}{K_i}c, K_i),\\
&{CDF}_{\nu_i}(V_i)  = \Pr(\nu_i \leq V_i),
\end{align*}
where
\begin{itemize}
\item[$V_i = $] $\lfloor{T_i - \alpha_i}\rfloor$, the maximum number of additional ``yes'' votes $X_i$ could receive such that it would remain below threshold.,
\item[$c = $] $10$, a hyperparameter chosen by the community based on their relative concern about false alarms versus (temporarily) missed detection.
\end{itemize}

@@

Using Eqs.~\eqref{eq3} and \eqref{eq5}, we can numerically compute the credibility of a vote's outcome in two key scenarios relevant for democratic communities.  For example, a community may only consider accepting a proposal when a vote's credibility is above $90\%$.

@@

To gain an intuitive grasp of the decision credibility used in a statistical quorum, we can examine a few toy scenarios:

In Figure \ref{fig:credible_interval}, we examine how the $90\%$ credible interval converges as more votes are collected out of a population of $100$.  A statistical quorum is achieved when the credible interval is contained entirely above zero, corresponding to a positive bias toward the proposal.  In this example, the latent bias of the community is that $75\%$ of the community supports the proposal while $25\%$ opposes it.  After approximately $35$ members have voted, a statistical quorum is achieved.

In Figure \ref{fig:simple_majority}, we examine how many affirmative votes are required from a community of size $100$ to reach a statistical quorum \textit{after} the proposal starts off with $10$ votes against it.  Initially, the $0-10$ vote tally in favor of rejecting the proposal gives the outcome over an $80\%$ credibility; however, as affirmative votes begin arriving, the credibility decreases until the `yes' votes outnumber the `no' votes.  After $21$ `yes' votes have been cast against the $10$ `no' votes, a statistical quorum of $90\%$ credibility is achieved.

In Figure \ref{fig:competing_proposals}, we examine how a set of three competing proposals gets resolved.  We simulate three proposals competing against each other in a community of size $100$, where the community has a latent bias of $75\%$, $50\%$, $25\%$ toward each of the proposals.  This scenario models what happens when a community slightly prefers the first proposal, is relatively indifferent or equivalently split on the second proposal, and slightly disproves of the third proposal.  Each member of the community casts a vote of $+0.5$, $0$, $-0.5$ for proposal 1, 2, 3, respectively.  (We have removed the stochasticity of the votes cast according to the community's latent biases in order to simplify the simulation.)  After approximately $35$ votes, we see that a statistical quorum of $90\%$ is achieved, signifying that the community's preference for proposal 1 over proposals 2, 3 is clear.

\begin{figure}[ht]
\centering
\includegraphics[width=0.6\textwidth]{figures/credible_interval.pdf}
\caption{Convergence of the $90\%$ credible interval as more votes are cast in the population of size $100$.  A statistical quorum is achieved when the credible interval is strictly above $0$.}
\label{fig:credible_interval}
\end{figure}

\begin{figure}[ht]
\centering
\begin{subfigure}[t]{.5\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/simple_majority.pdf}
\caption{Credibility of a proposal requiring a simple majority to pass: After $10$ `No' votes are initially cast, `Yes' votes arrive until a statistical quorum is achieved.}
\label{fig:simple_majority}
\end{subfigure}%
\begin{subfigure}[t]{.5\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/competing_proposals.pdf}
\caption{Credibility of a proposal competing against two other proposals: As more votes are cast according to the population's bias towards the three proposals ($75\%$, $50\%$, $25\%$), the first proposal eventually achieves a statistical quorum.}
\label{fig:competing_proposals}
\end{subfigure}
\caption{}
\end{figure}

\newpage
